---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
data<-read.csv("activity.csv")

```


## What is mean total number of steps taken per day?
```{r}
suppressMessages( library(dplyr))
data2<-group_by(data,date)
totalperday<-suppressMessages(summarize(data2,steps=sum(steps,na.rm = TRUE)))
totalperday<-as.data.frame(totalperday)
knitr::kable(totalperday,format = "html")
hist(totalperday$steps, main = "Histogram of total steps per day", xlab = "Steps")
meants<-mean(totalperday$steps)
meants<-round(meants,2)
mediants<-median(totalperday$steps)
```
The mean of the total number of steps per day is: *`r meants`* steps  
The median of the total number of steps per day is: *`r mediants`* steps 



## What is the average daily activity pattern?
```{r}
data3<-group_by(data,interval)
daily<-suppressMessages(summarize(data3, steps=mean(steps, na.rm = TRUE)))
plot(daily$interval,daily$steps, type = "l", xlab="interval", ylab = "steps")
winter<-which.max(daily$steps)
winter<-daily$interval[winter]
```
The maximum average steps per day are located in the interval *`r winter`*



## Imputing missing values
```{r cache=TRUE}
totalmis<-sum(is.na(data$steps))

data4<-data
#change the NA for the mean of that interval
for (i in 1:nrow(data4)) {
        if (is.na(data4[i,"steps"])==TRUE) {
                lin<-data4[i,"interval"]
                pos<-match(lin,daily$interval)
                data4[i,"steps"]<-daily[pos,"steps"]
        }
  }
#data4 is the new dataset without NA's
  newdata<-group_by(data4,date)
  newdata<-suppressMessages(summarize(newdata,steps=sum(steps)))
  hist(newdata$steps, main = "Histogram of total steps per day (imputed)", xlab = "steps")
  meanna<-mean(newdata$steps)
  meanna<-round(meanna,2)
  medianna<-median(newdata$steps)
  medianna<-round(medianna,2)

```
The total NAs present in the data are: *`r totalmis`*  

Replacing the NAs values with the average of steps in the corresponding interval leads to an increase in both the mean and median. The value for the mean is *`r toString(meanna)`*. The value for the median is *`r toString(medianna)`*. As can be seen both the mean and the median get the same value with this impute.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(chron)
data4$Wdays<-factor(chron::is.weekend(data4$date),levels = c(FALSE,TRUE), labels = c("Weekday","Weekend"))
library(ggplot2)
wdata<-group_by(data4,interval,Wdays)
wdata<-suppressMessages(summarize(wdata,steps=mean(steps)))
qplot(data=wdata,interval,steps, geom = "line",facets = Wdays~.)
```
  
    
It can be seen that in weekends are more activity in the evening, while in weekdays are more activity during the morning.



